Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
put in Josh 's introduction

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence.  Recent work in program induction show
that deep networks can be very successful at learning to write certain
kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
However, the success of these approaches -- both symbolic and neural --
relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
The DSL is an inventory of restricted programming primitives,
and imparts domain specific knowledge about the space of programs.
To what extent can we dispense with hand-engineered DSLs?

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \systemEnding,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to expose the commonalities across the programs that solve tasks within a domain,
letting the agent solve other related program learning problems.

>>>>

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill},
DeepCoder:~\cite{balog2016deepcoder}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain Specific Language} (DSL). The DSL is an
inventory of restricted programming primitives, encoding domain
specific knowledge about the space of programs to be considered. In
practice
we are often given only one or a few examples of each
program to be induced, and thus success often hinges on having a good
DSL, to
provide a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered DSLs?

We propose \emph{learning} the DSL. We consider the setting where we
have a collection of related programming tasks, each specified by a
set of input/output examples. We do not assume that the tasks are
annotated with ground-truth programs. We typically will not try to
learn a DSL completely from scratch, but rather to start from a weaker
or more general set of primitives in a Lisp-like representation and
construct a richer, more powerful and better-tuned domain-specific
language.

Our algorithm is called \system because it is based on a
novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
between ``wake'' and ``sleep'' phases to achieve three goals: finding
programs that solve each of the tasks; creating a DSL by discovering
and reusing domain-specific subroutines; and learning a neural network
that efficiently guides search for programs in the DSL. The DSL
distills commonalities across programs that solve the tasks, letting
the agent solve other related program learning problems. The neural
network ensures that searching for program solutions remains tractable
even as the DSL (and hence the search space for programs) expands.

%% Automatically inducing programs from examples is a long-standing goal
%% of artificial intelligence.  Recent work in program induction show
%% that deep networks can be very successful at learning to write certain
%% kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
%% DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
%% techniques (e.g., Metagol:~\cite{muggleton2015meta},
%% FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
%% However, the success of these approaches -- both symbolic and neural --
%% relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
%% The DSL is an inventory of restricted programming primitives,
%% and imparts domain specific knowledge about the space of programs.
%% To what extent can we dispense with hand-engineered DSLs?

%% This work proposes \emph{learning} the DSL.
%% We consider the setting where we have a collection of related programming tasks,
%% each specified by a set of input/output examples.
%% We do not assume that the tasks are annotated with
%% ground-truth programs.
%% Our algorithm, called \systemEnding,
%% does three things:
%% it finds programs that solve each of the tasks;
%% it creates its own DSL by discovering and reusing
%% domain-specific subroutines;
%% and it learns a neural network that
%% efficiently searches for programs in the DSL.
%% The purpose of the DSL
%% is to expose the commonalities across the programs that solve tasks within a domain,
%% letting the agent solve other related program learning problems.
